Supplemental Materials: \emph{Simultaneous segmentation and registration of diffusion MR images of the brain driven by active-contours}

### 2015-03-20 Publications/2015-NeuroImage/02-abstract.tex
fix aggregate, update results

% -*- root: 00-main.tex -*-
We propose a simultaneous segmentation and registration method called \emph{regseg}
  that uses active contours without edges extracted from structural images.
The contours evolve through a free-deformation field of B-Spline functions that
  optimally maps the surfaces onto the data in the target space.
\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
  in echo-planar imaging.
We proof the concept behind \emph{regseg} on four digital phantoms warped with random
  deformations, and conclude that the precision of the segmentation is subvoxel.
Finally, we demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
  datasets from the \acrlong*{hcp}.
The results suggest that \emph{regseg} overperformed an extended competing method.
\emph{Regseg} showed a consistent error within the 95\% confidence interval of
  $0.39-0.93mm$ in images of $1.25mm$ isotropic resolution.

>>>>

% -*- root: 00-main.tex -*-
We propose a simultaneous segmentation and registration method called \emph{regseg}
  that uses active contours without edges extracted from structural images.
The contours evolve through a free-deformation field of B-Spline functions that
  optimally maps the surfaces onto the data in the target space.
\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
  in echo-planar imaging.
We proof the concept behind \emph{regseg} on four digital phantoms warped with random
  deformations, and conclude that the precision of the segmentation is subvoxel.
Finally, we demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
  datasets from the \acrlong*{hcp}.
The results suggest that \emph{regseg} overperformed an extended competing method.
We compute the geometrical disagreement of the surfaces registered with \emph{regseg} with
  respect to a ground truth, obtaining a 95\% confidence interval of 0.56 - 0.66 (mm)
  for the error in images of 1.25mm isotropic resolution.

### 2015-03-20 Publications/2015-NeuroImage/06-discussion.tex
fix aggregate, update results

\paragraph*{Accuracy tests}
The hypothesis underpinning our research was that image registration can be reliably performed
  by searching for homogeneous regions in the target image that correspond to precise contours
  from an atlas, or extracted from other image (i.e. a different time step).
We proved that active contours without edges can be used to successfully drive a
  deformation supported by B-Spline functions using digital phantoms.
We randomly deformed 4 different phantom models mimicking three homogeneous regions,
  showing contrasts generated by a publicly available \gls*{mri} simulator.
After registration with \emph{regseg} we measured the Hausdorff distance between the
  projected contours using the ground-truth warping and our estimation.
We concluded that errors are significantly lower than voxel resolution.
We also assessed the 95\% \gls*{ci}, yielding an aggregated interval of
  $[0.64 - 0.66]mm$ in the low resolution phantoms ($2.0mm$ isotropic voxel) and
  $[0.34 - 0.38]mm$ in the high resolution phantoms ($1.0mm$ isotropic).
Therefore, we also conclude that the error is bounded and below the half of the
  voxel spacing to illustrate the accuracy of \emph{regseg}.

@@

After a preliminary experiment to demonstrate the aptness of our segmentation model
  in section \autoref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
  a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.
We processed 16 subjects from the \gls*{hcp} database with both \emph{regseg}
  and an in-house replication of the \gls*{t2b} correction method.
\emph{Regseg} showed very high accuracy, with an aggregated 95\% \gls*{ci} of $[1.08-1.50]mm$,
  almost below the pixel size of $1.25mm$.
The misregistration error remaining after \emph{regseg} was significantly lower ($p < 0.01$) than the
  error corresponding to the \gls*{t2b} correction, using the one-way ANOVA test
  (\autoref{tab:results_real}).
Visual inspection of all the results (\suppl{section S5}), and the violin plots included in
  \autoref{fig:results_real}, confirmed this intuition.
Even though we carefully configured the \gls*{t2b} method by using the same algorithm and
  same settings from a widely used software, cross-comparison experiments are prone to
  the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
Therefore, these results do not show that the \emph{regseg} software \emph{is} better than the
  \gls*{t2b} method, but they are communicated to proof that \emph{regseg} is a solid option to
  correct \gls*{dmri} data for distortions, and any other potential application.

>>>>

\paragraph*{Accuracy tests}
The hypothesis underpinning our research was that image registration can be reliably performed
  by searching for homogeneous regions in the target image that correspond to precise contours
  from an atlas, or extracted from other image (i.e. a different time step).
We proved that active contours without edges can be used to successfully drive a
  deformation supported by B-Spline functions using digital phantoms.
We randomly deformed 4 different phantom models mimicking three homogeneous regions,
  showing contrasts generated by a publicly available \gls*{mri} simulator.
After registration with \emph{regseg} we measured the Hausdorff distance between the
  projected contours using the ground-truth warping and our estimation.
We concluded that errors are significantly lower than voxel resolution.
We also assessed the 95\% \gls*{ci}, yielding an aggregate interval of
  $[0.64 - 0.66]mm$ in the low resolution phantoms ($2.0mm$ isotropic voxel) and
  $[0.34 - 0.38]mm$ in the high resolution phantoms ($1.0mm$ isotropic).
Therefore, we also conclude that the error is bounded and below the half of the
  voxel spacing to illustrate the accuracy of \emph{regseg}.

@@

After a preliminary experiment to demonstrate the aptness of our segmentation model
  in section \autoref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
  a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.
We processed 16 subjects from the \gls*{hcp} database with both \emph{regseg}
  and an in-house replication of the \gls*{t2b} correction method.
\emph{Regseg} showed very high accuracy, with an aggregate 95\% \gls*{ci} of $[1.08-1.50]mm$,
  almost below the pixel size of $1.25mm$.
The misregistration error remaining after \emph{regseg} was significantly lower ($p < 0.01$) than the
  error corresponding to the \gls*{t2b} correction, using the one-way ANOVA test
  (\autoref{tab:results_real}).
Visual inspection of all the results (\suppl{section S5}), and the violin plots included in
  \autoref{fig:results_real}, confirmed this intuition.
Even though we carefully configured the \gls*{t2b} method by using the same algorithm and
  same settings from a widely used software, cross-comparison experiments are prone to
  the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
Therefore, these results do not show that the \emph{regseg} software \emph{is} better than the
  \gls*{t2b} method, but they are communicated to proof that \emph{regseg} is a solid option to
  correct \gls*{dmri} data for distortions, and any other potential application.
